---
title: "Week 8 -- Confidence Intervals"
author: "Shane Hutton"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

This week we show how to build confidence intervals in R for t-tests
There is actually nothing new here because by default t.test() gives the CI

```{r}
OCD <- c(20,60,48,92,50,82,48,90,30,68,43,54,60,62,94,67,63,85)

#One-sample t-test, given population Mu of 77.43

t.test(OCD, mu=77.43, conf.level=.95)
```

Interpretation: We are 95% confident that the CI [51.59, 72.41] contains the true population mean. That is, over repeated sampling, 95 out of 100 confidence intervals will contain the population mean. 

Decision: Using the CI method we would reject the H0 because our population mu (77.43) is outside our interval [51.59, 72.41].

Note: We interpret the CI whether or not we reject/fail to reject (interpretation of the CI doesn't affect rejection)

Independent samples t-test

```{r}
calories<-data.frame(Speed=c("Slow","Slow","Slow","Slow","Slow","Slow","Fast","Fast","Fast","Fast","Fast","Fast"),Calories=c(700,450,850,600,450,550,450,800,750,700,550,650))

calories
t.test(calories$Calories~calories$Speed, var.equal=T)
```

Interpretation of the CI: We are 95% confident that the CI [-134.19, 234.19] contains the true mean difference. That is, over repeated sampling, 95 out of 100 confidence intervals will contain the population mean difference. 

Decision: We would fail to reject because the CI contains 0. 

Interpretation of the t-test: An independent samples t-test revealed no statistically significant difference between the fast (M=650, SD=130.38) and slow (M=600, SD=154.92) eaters on calorie consumption; t(10)=0.60, p>0.05.

Related samples t-test

```{r}
grades<-data.frame(Pretest=c(68,74,55,81,72,59,78,65,80),
                     Posttest=c(78,79,68,92,72,65,74,71,90))

t.test(grades$Posttest,grades$Pretest,paired=T)
```

Interpretation: We are 95% confident that the true mean difference lies within the CI [2.11, 10.56]. That is, over repeated sampling, 95 out of 100 confidence intervals will contain the population mean difference. 

Decision: We would reject the null, because the CI [2.11, 10.56] did not contain 0.

We found significance so we can calculate cohen's d

```{r}
#Dbar/SDdiff (recall how we found these last week)
6.33/5.5 #d = 1.15 large effect
```

Now we will calculate means and standard deviations for the interpretation

```{r}
mean(grades$Pretest)
sd(grades$Pretest)
mean(grades$Posttest)
sd(grades$Posttest)
```

Interpretation of t-test: A related samples t-test revealed a statistically significant difference between pre-test (M=70.22, SD=9.22) and post-test (M=76.56, SD=9.30) grades; t(8)=3.45, p<.05, d=1.15. Cohen's d signified a large effect.