---
title: "Week 7 -- Independent samples and related samples t-tests"
author: "Shane Hutton"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

This week we learn how to conduct two-sample t-tests in R

We will start with independent samples t-test

Let's look at sample data given for example 9.2 (p. 297) in the book
The data: the number of calories consumed in a meal between groups asked to eat slowly or fast

```{r}
calories <- data.frame(Slow=c(700,450,850,600,450,550), Fast=c(450,800,750,700,550,650))

mean(calories$Slow)
mean(calories$Fast)

t.test(calories$Fast, calories$Slow, var.equal=T)
```

What about  using the formulas learned in async?
Calculating the t-statistic "by hand" (I have created the formula for you, all you need to do is type in the values)
First we will get the mean and sd for each group

```{r}
mean(calories$Slow)
mean(calories$Fast)
sd(calories$Slow)
sd(calories$Fast)
```

Then we will plug them into the objects and run the code to obtain the t-value using the formula from async

```{r}
M1 <- 650 #sample mean of fast
M2 <- 600 #sample mean of slow
s1 <- 130.384 #sd of fast
s2 <- 154.919 #sd of slow
n1 <- 6 #sample size of fast
n2 <- 6 #sample size of slow
t <- (M1-M2)/sqrt((((n1-1)*s1^2+(n2-1)*s2^2)/(n1+n2-2))*((1/n1)+(1/n2)))
t
```

Calculating the p-value
We still have to multiply the probability for two-tailed tests (note: the t-table makes this easier)
We still need to be careful with the lower.tail command (because I have a positive t-statistic I am using FALSE)
remember df = 6 + 6 - 2 = 10

```{r}
2*pt(t, df=10, lower.tail=FALSE)
```

Calculating effect size -- note, we're doing it here for demonstration, but given our results above, we would not calculate effect size.
Here again, I created the formula for you but for these to work the objects above have to be defined

```{r}
#Calculate Cohen's d
(M1-M2)/sqrt((((n1-1)*s1^2+(n2-1)*s2^2)/(n1+n2-2)))

#Calculate eta squared
(t^2)/(t^2+(n1 + n2 - 2))
```

Conclusion: An independent-samples t-test revealed no statistically significant difference between the number of calories consumed between the fast eaters (M=650, SD=130.38) and the slow eaters (M=600, SD=154.92), t(10)=0.60, p>.05. 

Now, lets run code for related samples t-test

```{r}
grades <- data.frame(Pretest=c(68,74,55,81,72,59,78,65,80), 
                   Posttest=c(78,79,68,92,72,65,74,71,90))

t.test(grades$Posttest, grades$Pretest, paired=TRUE)
```

Now run this "by hand"

```{r}
#First we will create the difference scores (i.e., Posttest - Pretest) 

grades$difference <- c(grades$Posttest - grades$Pretest)

#Look at the data and see the difference scores
grades

#Now we will find the mean and sd of the difference scores and use them to compute t using the formula from async
mean(grades$difference) #this is Dbar
sd(grades$difference)

teststatforrelated <- 6.33/(5.5/sqrt(9))

#now, let's find the p-value
2*pt(teststatforrelated, df=8, lower.tail=FALSE)

```

Let's calculate cohen's d, because we found statistical significance
#Dbar/SDdiff

```{r}
6.33/5.5 

#d = 1.15 large effect
```

Find means for APA write-up/conclusion

```{r}
mean(grades$Pretest)
sd(grades$Pretest)
mean(grades$Posttest)
sd(grades$Posttest)
```

Conclusion: A related-samples t-test revealed a statistically significant difference between the pre-test (M=70.22, SD=9.22) and post-test (M=76.56, SD=9.30); t(8)=3.45, p<.05, d=1.15. Cohen's d revealed a large effect.
