---
title: "Week 10 - Two-Way ANOVAs"
author: "Shane Hutton"
output:
  pdf_document: default
---

This week we will learn how to conduct a two-way ANOVA in R

First, set the working directory
Let's look at the sleep data from async
```{r}
sleep <- read.table("sleepdata.txt", header=T, sep="\t")
```

Now, let's run the two-way ANOVA
```{r}
sleep.aov <- aov(Sleep~Timing + Intensity + Timing:Intensity, data=sleep)
summary(sleep.aov)
```

Let's create an interaction plot using R
```{r}
interaction.plot(sleep$Intensity, sleep$Timing,  sleep$Sleep,
                 ylim=range(15,45),xlab="Intensity",
                 ylab="Sleep",
                 col=c("red","blue"), main="Plot of Timing x Intensity")
```

There is a problem because R is ordering Low, Medium and High in alphabetical order but that doesn't make sense
However, we can force R to create it how we want it -- Low, Medium and High by using the reorder function
```{r}
reorder <- factor(sleep$Intensity, levels=c("Low", "Medium", "High"))
```

Now, we will modify the interaction.plot code to have reordered Intensity in place of where we originally have intensity 
```{r}
interaction.plot(reorder, sleep$Timing,  sleep$Sleep,
                 ylim=range(15,45), xlab="Intensity",
                 ylab="Sleep",
                 col=c("red","blue"), main="Plot of Timing x Intensity")
```