---
title: "Week 14 - Chi-Square"
author: "Shane Hutton"
output:
  pdf_document: default
---

This week we will learn how to conduct a chi-square test in R

We will start with the goodness-of-fit test (only one discrete var)

Let's look at the test format data from async
```{r}
format <- c(17,54,64)
```

We will run the chi-square test using chisq.test
We need to plug in the expected proportions for each group (all equally weighted in this example) three groups - so we have 1/3 expected in each group
```{r}
chisq.test(format, p=c(1/3, 1/3, 1/3))
#Note: unless you're given data for the expected frequencies, use an equal distribution of data
```

Now, let's look at the test for independence (is there an association between two discrete variables?)

First, we will import the data after setting the working directory
```{r}
graduate <- read.table("graduate.txt", header=TRUE, sep="\t")
```

label the levels of the variables for the table we're building (this is optional - but helfpul to read your table)
```{r}
graduate$Business <- factor(graduate$Business, levels= c("0","1"), labels = c("Non-bus", "Bus"))
graduate$Doctorate <- factor(graduate$Doctorate, levels= c("0","1"), labels = c("Masters", "Doctorate"))
```

Now, let's summarize the data in a table (call it gradtable)
```{r}
gradtable <- table(graduate$Business, graduate$Doctorate) 
#look at the table after you create it
gradtable
```

Run the chi-square test
```{r}
chisq.test(gradtable, correct=F)
#Note: we use correct = F to get the same test statistic we computed by hand (no continuity correction)
#for some future analyses you may want the continuity correction/you should do research to determine when this may be the case
```

Or we could input the values into the table
```{r}
gradtable2 <- data.frame(c(574,176), c(185,3))

chisq.test(gradtable2, correct=F)
```

Based upon using a chi-square test for independence, we determined that there was a statistically significant association between whether you have a business degree and whether you pursue a masters or a doctoral degree, X2(1)=46.57, p<0.05.

```{r}
#calculating odds ratio for odds of conferring masters for biz vs. non-bus majors
(185/3)/(574/176) 
#odds of pursuing a masters degree in business compared to pursuing a masters degree in non-business. 
#odds ratio=18.91
```

The odds of pursuing a masters degree in business are 18.91 times higher than the odds of pursuing a masters degree in non-business. 