---
title: "Week 12 - SLR"
author: "Shane Hutton"
output:
  pdf_document: default
---

This week we will learn how to conduct a SLR (simple linear regression) in R

Let's look at the phd data from async
First set your working directory

Now we will import the data
```{r}
phd <- read.table("phd.txt", header=TRUE, sep="\t")
phd
```

Now we will make a scatterplot to examine the data visually
```{r}
plot(phd$Time, phd$Pubs ,
     main="Scatterplot of Time & Pubs", 
     xlab="Time Since Grad (in Years)", ylab="Publications", pch=19, 
     xlim=c(0,20), ylim=c(0,50))
```

Let's run an SLR with Pubs as DV and Time as IV; we will save it to an object called phd.regression
REGRESSIONOBJECTNAME <- lm(YVARIABLE ~ XVARIABLE, data=DATASETNAME)
```{r}
phd.regression <- lm(Pubs ~ Time, data=phd)
```

Here are the results:
```{r}
#Parameter estimates
phd.regression

#More information for hypothesis tests
summary(phd.regression)

#confidence intervals for parameter estimates
confint(phd.regression)

range(phd$Pubs)
```

Plotting regression line on scatterplot
```{r}
plot(phd$Time, phd$Pubs ,
     main="Scatterplot of Time & Pubs", 
     xlab="Time Since Grad (in Years)", ylab="Publications", pch=19, 
     xlim=c(0,20), ylim=c(0,50))

abline(phd.regression, col="blue")
```

An additional example (using data that is stored in R)
The dataset in R is called mtcars
```{r}
mtcars
```

Let's examine a possible relationship between car weight & miles per gallon
```{r}
plot(mtcars$wt, mtcars$mpg, 
     main="Scatterplot of Weight and MPG", 
     xlab="Weight in 1000s", ylab="MPG", pch=19)
```

Does car weight predict MPG? 
```{r}
cars.regression <- lm(mpg ~ wt, data=mtcars)
summary(cars.regression)
```