---
title: 'Applied Statistics - Week 1: Introduction to R'
author: "Shane Hutton"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

## Welcome to R! 
Throughout this term, we will use R Markdown (or .Rmd) files to keep
our weekly live session code and assignment code organized. Today, we'll start with our
first .Rmd file. 

Here, you can enter text that you want to print out in your final output file (just like in a Word document). Next, we're going to see where we can put our code... in things called "code chunks." These chunks tell R to interpret what's in them as code (rather than text).

```{r}
2+2
```


```{r}
#You can also include text notes in code chunks; however, to let R know 
#that they are text and not code, they need to begin with the # symbol.

#First we will see how R can perform simple calculations
#Suppose we want to calculate 2+2
2+2
3+3 #you can add notes anywhere
```

What about the square root of 9?
```{r}
sqrt(9) #sqrt is an example of a built-in R function - we'll learn a lot 
#of these this semester
```

Additional calculations using R as a calculator
```{r}
9-2

9^2 #the carrot ^ symbol is used to exponentiate

3*6 #asterisk * is used for multiplication
```

## Defining objects in R

R stores things (numbers, data, output, etc.) as objects. We will learn this term how to store the findings from statistical analyses as objects, but let's start with a simpler case.

We can assign values to objects (let's call them x and y) and then do the calculations 
with those objects. This practice is especially handy if you're going to want to use thesame number (or object) in numerous calculations.

To assign an object in R, you choose a name for that object and then tell R what the object is equal to. 
```{r}
#let's assign the number 9 to an object named x
x <- 9 #we can do this using a backwards arrow

#you can also assign objects using the equals sign
z = 4 

y <- 2 #this assigns the number 2 to y

#Note: you don't have to choose single-letter object names
whatever <- 52

x
y
z
whatever
```

Futhermore, you can see the objects in the environment window to the right or you can "print" them out by writing the name of the object in the R code.

```{r}
#print the object x
x

#print the object whatever
whatever

```

Let's use some of our stored objects in calculations.
```{r}

x

y+z 

sqrt(x)

whatever - y

z^y
```

Next, we will learn how to create vectors and dataframes and store those as objects in R. 

To create a vector of elements, we use the code: c(element, element, element,...)
```{r}
#Let's create a vector with the following numbers: 4 5 63 67 89
vector1 <- c(4,5,63,67,89) 

#Let's create another vector with the numbers: 88 92 71 69 99
vector2 <- c(88, 92, 71, 69, 99)

#now let's create a vector with string (or non-numeric) data: dog, cat, bunny
vectorstring <- c("dog", "cat", "bunny") #for non-numeric data, we use "" to 
#show R that these aren't numbers
```

R will will also store the vectors in the environment window or you can print them.
```{r}
#print the string vector
vectorstring

#print the numeric vector
vector1
```

Sometimes, we have two vectors that are related (i.e., they are from the same dataset)
and we want to put them together in a single data frame. We can do this using another 
built-in R function.
```{r}
#Let's make a dataframe out of our two numeric vectors. We can do this 
#because they're both the same size (5 elements each). I'm going to store 
#the dataframe in an object.
week1dataframe <- data.frame(vector1, vector2)

#print the dataframe
week1dataframe
```

data.frame is a function -- we will be using a lot of functions. If you need help with 
a function you can just type help() with the name of the function in the ()
```{r}
help(data.frame)
```

Suppose we want to change the names of our vectors within in our dataframe so that
they are meaningful and tell us something about the data. For the vectors we have, 
let's say that vector 1 is the student ID and vector 2 is the student's last grade on 
an exam.
```{r}
names(week1dataframe) <- c("SID", "Score")

week1dataframe
```

Let's say we also have data on student sex and want to add that to our existing dataframe. We can also do that. We're going to use the $ symbol after our dataframe name to let R know that we want to attach the new vector to the existing data. 
```{r}
#let's add a new variable: Sex
week1dataframe$Sex <- c("M","F","F","M","F")

week1dataframe
```

```{r}
#let's add another variable: age
week1dataframe$Age = c(29,20,25,27,35)

week1dataframe
```

If we just want to view the values from one variable in a dataframe, you can use 
call attention to just one variable (or column) with that same $ symbol.
```{r}
#let's just look at the scores data
week1dataframe$Score
```

## Frequency Distributions

Next, we're going to learn how to create an ungrouped frequency distribution

First, we will create a vector from the math ability data in async.
```{r}
mathability <- c(6,5,6,1,4,4,5,6,5,3,5,6,7,3,5,5,4,5,4,7,2,5,5,2,6,5,5,4,5,3,5,7,6,5,3)
mathability
```

Then, we're going to use R's built in "table" function to create a frequency table.
```{r}
table(mathability)

#Let's make the table look a little better using transform - this is 
#like the transpose function in excel
transform(table(mathability))
```

Now, let's add some information to the table (percent, cumulative frequency and cumulative percent)
```{r}
freq_table <- transform(table(mathability)) #save the table to an object
n <- length(mathability) #store the total number of elements in the original 
#vector in the object n

#build the table - adding relative frequency, cumulative frequency, and 
#cumulative percent
transform(freq_table, Rel_Freq=prop.table(Freq)*100, Cum_Freq=cumsum(Freq), Cum_Percent=(cumsum(Freq)/n)*100)
```

In addition to building tables and doing computations incredibly fast, R is also 
great at visual depictions of data. Let's start with a very simple example for this week and let's build a histogram. Let's use the Cash data from async.
```{r}
#First, enter in the data as a vector
cash <- c(.05,6,30,0,65,4.75,50,0,100,.17,15,210,2,175,42,65,2,
          200,0,10,0,15,20,30,
          12,0,30,45,7,10,80,12,28,.75,100,33,1,14,5,23,45)

#To create the histogram, we'll use the R function hist
hist(cash)

help(hist)
```

As the final step - you can "knit" this file to a pdf, word document or HTML.