---
title: "Week 5 - One-sample t-test"
author: "Shane Hutton"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

Let's use R to perform calculations for couples example.

First, let's calculate the t-statistic "by hand"
```{r}
(2.6-3)/(1.9/sqrt(70))
```

Next, we can calculate the p-value.
Instead of pnorm, we will use pt, the only difference is we have to specify the df
We still have to multiply the probability for two-tailed tests (note: the t-table makes this easier).
We still need to be careful with the lower.tail command (because we have a negative t-statistic we need to use TRUE).

```{r}
2*pt(-1.76,df=69, lower.tail=TRUE)

#or we could have used the positive value (symmetry)

2*pt(1.76,df=69, lower.tail=FALSE)
```

Now, let's learn how to conduct a t-test when given a dataset.
Let's look at sample data given for example 9.1 in the book.

```{r}
OCD <- c(20,60,48,92,50,82,48,90,30,68,43,54,60,62,94,67,63,85)

OCD

t.test(OCD,mu=77.43, conf.level=.95)
```

We can also use R to calculate the effect size "by hand"

```{r}
#Calculate Cohen's d:
(62-77.43)/20.94
```
