---
title: "Week 4 - Normal Probabilities in Sampling Distributions"
author: "Shane Hutton"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

This week we will learn how to compute normal probabilities in sampling distributions. 

Let's first use R as a calculator to compute total possible samples for theoretical sampling and then for experimental sampling.

How many samples of four participants (n = 4) from a population of ten people (N = 10) would we have using theoretical sampling?
```{r}
#The formula here is: N^n
10^4
```

How many samples of four participants (n = 4) from a population of ten people (N = 10) would we have using experimental sampling?
```{r}
#The formula here is: N!/n!(N-n)!
factorial(10)/(factorial(4)*factorial(10-4))
```

### Recall the IQ test example from last week. 

IQ tests are normed to have a mean of 100 and standard deviation (sd) of 15. Last week, we found the probability that an individual scores greater than 120 using two different methods.
```{r}
#Method 1 - using the z-score in pnorm
(120-100)/15 #this finds the z-score
pnorm(1.33, lower.tail=FALSE)

#Method 2 - using the raw score in pnorm
pnorm(120, mean=100, sd=15, lower.tail=FALSE)
```

### This week, we're going to work with a sampling distribution. 

IQ tests are normed to have a mean of 100 and standard deviation of 15. What is the probability that a sample of 9 individuals has a mean greater than 110 on an IQ test?

First, we need to compute the standard error (se), then the z-score using the standard error.
```{r}
#standard error formula = sigma/sqrt(n)
15/sqrt(9)

#find the z-score using the se 
(110-100)/5

#then use pnorm just like before
pnorm(2, lower.tail=FALSE)
```

Interpretation: There is as 2.28% probability that a sample of 9 individuals has a mean greater than 110 on this IQ test.

Just like last week, we could have skipped computing the z-score and coded in the mean and se
```{r}
#note: the code still says "sd" even though, we're using the standard error and 
#that's ok... because the se is the sd of the sampling distribution. It is up to you
#to put the se in here, though, and not the sd, when working with sampling #distributions. 
pnorm(110, mean=100, sd=5, lower.tail=FALSE)
```

Also, we could have done this whole thing in one step (calculate the se within the code)
```{r}
pnorm(110, mean=100, sd=15/sqrt(9), lower.tail=F)
```

Example: 
Suppose a standardized test is normed to have a mean of 140 and standard deviation of 12.

What is the probability that a sample of 5 individuals has a mean less than 130 on the standardized test?
```{r}
pnorm(130, mean=140, sd=12/sqrt(5), lower.tail=TRUE)
```
Interpretation: The probability of a sample of 5 individuals having a mean less than 130 is 3.12%.

What's the probability that a sample of 10 has a mean greater than 150 on the same standardized test?
```{r}
pnorm(150, mean=140, sd=12/sqrt(10), lower.tail=F)
```
Interpretation: The probability of a sample of 10 individuals having a mean greater than 150 is 0.42%.

Example:
Suppose we have state standardized test that's normed to have a mean of 170 and an sd of 15. 

What is the probability that a sample of 7 students has a mean greater than 175?
```{r}
pnorm(175, mean=170, sd=15/sqrt(7), lower.tail=F)
```
Interpretation: The probability of a sample of 7 students having a mean greater than 175 is 18.89%.

What is the probability that a sample of 17 students has a mean greater than 175?
```{r}
pnorm(175, mean=170, sd=15/sqrt(17), lower.tail=F)
```
Interpretation: The probability of a sample of 17 students having a mean greater than 175 is 8.47%.

What is the probability that a sample of 27 students has a mean greater than 175?
```{r}
pnorm(175, mean=170, sd=15/sqrt(27), lower.tail=F)
```
Interpretation: The probability of a sample of 27 students having a mean greater than 175 is 4.16%.










